---
title: "Homework 4"
author: "Minxuan Wang"
date: "November 2017"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
## a.
```{r}
data.cost<-read.table("TableF4-4.txt",head=TRUE)
attach(data.cost)

## Compute each terms
lnCPf<-log(cost/pf)
lnQ<-log(q)
lnQ2<-0.5*(lnQ)^2
lnPkPf<-log(pk/pf)
lnPlPf<-log(pl/pf)

## Linear regression
fitlm_cost<-lm(lnCPf~lnQ+lnQ2+lnPkPf+lnPlPf)
summary(fitlm_cost)

## Asymptotic covariance matrix
cov_matrix<-vcov(fitlm_cost)
cov_matrix
```

## b.
```{r}
## Compute delta f
delta_f<-1-coefficients(fitlm_cost)[4]-coefficients(fitlm_cost)[5]
delta_f
estmean<-coef(fitlm_cost)[4:5]
estvar<-vcov(fitlm_cost)[4:5,4:5]

## Estimate the asymptotic standard error
library("msm")
deltamethod(~1-x1-x2,estmean,estvar)
```

## c.
```{r}
beta<-coefficients(fitlm_cost)[2]
gamma<-coefficients(fitlm_cost)[3]
est.Q<-exp((1-beta)/gamma)
est.Q
est.Q_mean<-coef(fitlm_cost)[2:3]
est.Q_var<-vcov(fitlm_cost)[2:3,2:3]

## Standard error
se.Q<-deltamethod(~exp((1-x1)/x2),est.Q_mean,est.Q_var)
lowerbound<-est.Q-qnorm(0.975)*se.Q
upperbound<-est.Q+qnorm(0.975)*se.Q

## 95% confidence interval
IC<-c(lowerbound,upperbound)
IC 
```

## d.
```{r}
## Pick out the firms sets
firms1<-subset(data.cost,data.cost$q>=lowerbound)
firms2<-subset(firms1,firms1$q<=upperbound)

## Compute the number of firms that reached the efficient scale
length(firms2$q)
```

# Question 2
## a.
```{r}
setwd("D:/Econ 403A/Homework 4")
merged.data<-read.csv("Koop-Tobias.csv") # Get from NYU Stern

## Define the variables name in R 
educ<-merged.data$EDUC
logwage<-merged.data$LOGWAGE
potexper<-merged.data$POTEXPER
ability<-merged.data$ABILITY
mothered<-merged.data$MOTHERED
fathered<-merged.data$FATHERED
brknhome<-merged.data$BRKNHOME
siblings<-merged.data$SIBLINGS 

## Linear regression
lm1<-lm(logwage~educ+potexper+ability)
lm1
lm2<-lm(logwage~-1+mothered+fathered+brknhome+siblings)
lm2
```

## b/c.   
The F-test is the statistic for the hypothesis test with null hypothesis and alternate hypothesis:   
  H0: All non-constant coefficients in the regression equation are zero   
  Ha: At least one of the non-constant coefficients in the regression equation is non-zero.
```{r}
summary(lm1)
summary(lm2)
```
p-value: < 2.2e-16, which means that we reject the H0, model has predictive capability.   

Numerically, we can the defination of F statistic (using the first model as an example): 
```{r}
anova(lm1)
## SS(Regression)=SS(Total)-S(Residual)

## Get the SST value
SST_1<-var(logwage)*(nrow(merged.data)-1)

## Get the SSE value
SSE_1<-sum(lm1$residual^2)

## Get the SSR value
SSR_1<-SST_1-SSE_1

## Get the degree of freedom
dfE_1<-lm1$df.residual
dfReg_1<-nrow(merged.data)-1-dfE_1
MSreg_1<-SSR_1/dfReg_1
MSE_1<-SSE_1/dfE_1
Fstat_1<-MSreg_1/MSE_1
pvalue_1<-pf(Fstat_1,dfReg_1,dfE_1,lower.tail=FALSE)
```

## d. Wald test
```{r}
library(survey)
regTermTest(lm2,"mothered")
regTermTest(lm2,"fathered")
regTermTest(lm2,"brknhome")
regTermTest(lm2,"siblings")
```

# Question 3
## (i). 
```{r}
load("D:/Econ 403A/Homework 4/401ksubs.RData")
attach(data)
nettfa<-data$nettfa
mean(nettfa)
sd(nettfa)
max(nettfa)
min(nettfa)
```

## (ii).
```{r}
## T test
nettfa_0<-subset(nettfa,data$e401k==0)
nettfa_1<-subset(nettfa,data$e401k==1)
t.test(nettfa_0,nettfa_1,
alternative="two.side",
paired=FALSE,
var.equal=FALSE,
conf.level=.95)
```
From the result of t.test, p-value < 2.2e-16, which means we should reject the H0: the average nettfa does not differ by 401(k) eligibility status.    
The dollar amount difference is:
```{r}
mean(nettfa_1)-mean(nettfa_0)
```

## (iii). 
```{r}
e401k<-data$e401k
inc2<-incsq
age2<-agesq
fitlm=lm(nettfa~inc+inc2+age+age2+e401k)
summary(fitlm)
```
From the p-value, this regression model is statistically significant. So the linear equations are:   
when e401k=1, $nettfa=-23.24+0.008109\times incsq+0.01221\times agesq+8.166$   
when e401k=0, $nettfa=-23.24+0.008109\times incsq+0.01221\times agesq$  
The estimated dollar effect of 401(k) eligibility is 8.166    

## (iv). 
```{r}
age_41<-age-41

## Regress the model with interaction term
fitlm_2=lm(nettfa~inc+inc2+age+age2+I(e401k*age_41)+e401k)
summary(fitlm_2)
```
The interaction term is significant because p-value=1.55e-07(t=5.251), the coeffecient is 0.638.   

## (v).      
The coefficient on e401k at age 41 in these two regressions are 9.705(in part iii the influences at all ages are same) and 9.585, it doesn't differ a lot.   

## (vi).   
```{r}
## Define the dummy virables
fsize1<-as.numeric(fsize==1)
fsize2<-as.numeric(fsize==2)
fsize3<-as.numeric(fsize==3)
fsize4<-as.numeric(fsize==4)
fsize5<-as.numeric(fsize>=5)

## Add dummy virables to regression model
fitlm_3<-lm(nettfa~inc+inc2+age+age2+e401k+fsize2+fsize3+fsize4+fsize5)
summary(fitlm_3)
anova(fitlm_3)
```
## (vii).   
```{r}
## Define five conditions
data_fsize1<-subset.data.frame(data,fsize1=="1")
data_fsize2<-subset.data.frame(data,fsize2=="1")
data_fsize3<-subset.data.frame(data,fsize3=="1")
data_fsize4<-subset.data.frame(data,fsize4=="1")
data_fsize5<-subset.data.frame(data,fsize5=="1")

## Run the regression
unreg.1<-lm(nettfa~inc+incsq+age+agesq+e401k,data=data_fsize1)
unreg.2<-lm(nettfa~inc+incsq+age+agesq+e401k,data=data_fsize2)
unreg.3<-lm(nettfa~inc+incsq+age+agesq+e401k,data=data_fsize3)
unreg.4<-lm(nettfa~inc+incsq+age+agesq+e401k,data=data_fsize4)
unreg.5<-lm(nettfa~inc+incsq+age+agesq+e401k,data=data_fsize5)

## review the regression results
anova(unreg.1)
anova(unreg.2)
anova(unreg.3)
anova(unreg.4)
anova(unreg.5)
anova(fitlm_3)

## Calculate sum of squared residuals for each regression
SSR_ur<-sum(anova(unreg.1)[6,2],anova(unreg.2)[6,2],anova(unreg.3)[6,2],anova(unreg.4)[6,2],anova(unreg.5)[6,2])
SSR_ur
SSR_r<-anova(fitlm_3)[10,2]
SSR_r

## Computing the Chow test statistic (F-test)
Chow.F.statistic<-((SSR_r-SSR_ur)/SSR_ur)*(9245/20)
Chow.F.statistic

## Calculate P-value
1-pf(Chow.F.statistic,20,9245)
```
From the result we can see the p-value is essentially zero. In this case, there is strong evidence that the slopes change across family size.

# Question 4
## (i).   
```{r}
## Estimate simple linear probability model
fitlm_e401k=lm(e401k~inc+inc2+age+age2+male,data=data)
summary(fitlm_e401k)
library(RCurl)

## Import the function
url_robust<-"https://raw.githubusercontent.com/IsidoreBeautrelet/economictheoryblog/master/robust_summary.R"
eval(parse(text=getURL(url_robust,ssl.verifypeer=FALSE)),envir=.GlobalEnv)

## Use new summary function
summary(fitlm_e401k,robust=TRUE)
```
All the parameters are statistical significant:   
$$\widehat{e401k}=-0.506+0.0124\times inc-0.000062\times inc^2+0.0265\times age- 0.00031\times age^2-0.0035\times male$$
From the two summaries, we can see the Std.Error are almostly the same. So there are no important differences.

## (ii).   
Notice that the approximate estimator of the random error term $\mu_{i}$ is expressed by the residual $e_{i}$, such that we get:     
$$Var(\mu_{i})=E(\mu_{i}^2)\approx e_{i}^2$$   
$$e_{i}=Y_{i}-(\hat{Y_{i}})_{ols}$$   
We can write this as a regression model in a simple way.   
$$e_{i}^2=\alpha_0+\alpha_1X_i+\alpha_2X^2+v$$   
The restrictions are $\alpha_0=0$,$\alpha_1=1$, and $\alpha_2=-1$.
In the oringinal linear probability model:
$$\hat{Y_i}=\beta_{0}+\beta_{i}X_{i}$$
So, when we run the regression $e_{i}^2$ on $\hat{y}_i$ and $\hat{y_{i}^2}$, the intercept estimates should be close to zero, the coefficient on $\hat{y}_i$ should be close to 1, and the coefficient on $\hat{y_{i}^2}$ should be close to ¨C1.

## (iii).   
```{r}
## Get the residual squared sequence
u2<-fitlm_e401k$residuals^2

## Do the linear regression ui2 on yi and yi2
y<-fitted(fitlm_e401k)
summary(lm(u2~y+I(y^2)))
```
The White F statistic is about 310.3, which is very significant. The coefficient on
$\widehat{e401k}$ and $\widehat{e401k}^2$ is 1.0097 and -0.9703, the intercept is -0.009.
The coefficient estimates roughly correspond to the theoretical values described in part (ii).   

## (iv).   
```{r}
## Compute the upper bound and lower bound of fitted values
max(y)
min(y)

## Fit a WLS model using weights=1/(fitted values)
fitlm_e401k.wls=lm(e401k~inc+inc2+age+age2+male,data=data,weights=1/y)
summary(fitlm_e401k.wls)
```
$$\widehat{e401k}=-0.4841+0.01277\times inc-0.000062\times inc^2+0.025\times age-0.00029\times age^2-0.00324\times male$$
They doesn't differ in important ways from the OLS estimates.

# Question 5
## (i).   
```{r}
fitlm2<-lm(nettfa~inc+inc2+age+age2+male+e401k)
summary(fitlm2)
```
$$\widehat{nettfa}=21.198-0.27\times inc+0.0102\times inc^2-1.940\times age+0.0346\times age^2+3.369\times male+9.713\times e401k$$
The coefficient on e401k is 9.713, which means when other terms are fixed, the mean of net financial assets of a family with e401k=1 is about 9713 greater than the family with e401k=0.   

## (ii).   
Same as the previous question, we pick out the residuals first and then do the regression of $\hat{\mu_{i}}^2$ on $inc$, $inc^2$, $age$, $age^2$, $male$ and $e401k$.
```{r}
## Define the square of residuals
u2_2<-fitlm2$residuals^2

## Regress the linear model
summary(lm(u2_2~inc+inc2+age+age2+male+e401k))
```
$R^2$=0.0374, F-Statistic is 59.97, p-value: < 2.2e-16. So this model could have heteroskedasticity, which means given the explanatory variablees, the variance of series error is not equal to 0.

## (iii).   
```{r}
library(L1pack)
fitlad<-lad(nettfa~inc+inc2+age+age2+male+e401k)
summary(fitlad)
```
The lad estimate model is:
$$\widehat{nettfa}=12.4912-0.2616\times inc+0.0071\times inc^2-0.7227\times age+ 0.0111\times age^2+1.0188\times male+3.7373\times e401k$$
$\beta_6$ is the cofficient of e401k, $\beta$=3.7373, which means when other terms are fixed, the median net financial assets of a family whoes e401k=1 is about 3737 greater than the family with e401k=0.

## (iv).
401(k) eligibility has a larger effect on mean wealth than on median wealth, which means the 401(k) eligibility has a larger effect when net financial assets are in a high level.